---
title: "Random Forest Training Report"
author: "JW"
date: "`r Sys.Date()`"
output: pdf_document
params:
  errors: NULL
  wd: NULL
  data_dir: NULL
  output_dir: NULL
  session_info: NULL
  model_name: NULL
  start: NULL
  dataset: NULL
  dataset_out_dir: NULL
  skim: NULL
  metrics: NULL
  size: NULL
  end: NULL
  plots: NULL
  rf_files: NULL
  bspid: NULL
  subbspids: NULL
header-includes:
  - \usepackage{fvextra}
  - \RecustomVerbatimEnvironment{verbatim}{Verbatim}{breaklines=true}
  - \usepackage{pdfpages}
---

```{r Helpers, echo=FALSE}
# Shared plotting and PDF helper functions.
source("helpers.R")
```

```{r Import, echo=FALSE}
# Stack per-BSP metric tables and split BSP id into class columns.
metrics_df = dplyr::bind_rows(params$metrics, .id = "bsp_id", ) |> 
  tidyr::separate(bsp_id, into = c("class1","class2","class3"), sep = "-", fill = "right", remove = FALSE)
```

\newpage

## Session info

```{r SessionInfo, echo=FALSE, results='asis', tidy=FALSE}
si <- capture.output(params$session_info)
cat(strwrap(si, width = 90), sep = "\n")
```

\newpage

# Model training characteristics

Dataset: `r params$dataset`

Start: `r params$start`

End: `r params$end`

Directories:

```{r Directories, echo=FALSE}
data.frame(
  name = c("Working directory", "Data directory", "Output directory", "Dataset output directory"),
  dir = c(params$wd, params$data_dir, params$output_dir, params$dataset_out_dir)
) |> 
  knitr::kable(col.names = c("", ""), booktabs = TRUE) |>   
  kableExtra::kable_styling(position = "left")
```

\newpage

# Results

\section*{True Positive Rate}

The True Positive Rate (TPR), also called \emph{sensitivity}, measures how well a test identifies actual cases. It is calculated as the proportion of \emph{true positives} ($TP$) among all actual positive cases, i.e., \emph{true positives} plus \emph{false negatives} ($FN$):

\[
\text{TPR} = \frac{TP}{TP + FN}
\]

A high TPR means nearly all sick individuals are correctly detected, while a low TPR shows many cases are missed.

```{r True positive rate, echo=FALSE, fig.height=7}
metric_plot(
  m = metrics_df[,c("class", "n", "P", "TPR", "bsp_id", "class1")],
  metric_title = "True positive rate (TPR)",
  metric_shrt = "TPR [-]"
)
```

\newpage

\section*{Positive Predictive Value}

The Positive Predictive Value (PPV) indicates how likely it is that a person actually has a disease when the test result is positive. It is calculated as the proportion of \emph{true positives} ($TP$) among all positive test results, i.e., \emph{true positives} plus \emph{false positives} ($FP$):

\[
\text{PPV} = \frac{TP}{TP + FP}
\]

A high PPV means positive results are mostly correct, while a low PPV indicates many positives may be false.

```{r Positive predictive value, echo=FALSE, fig.height=7}
metric_plot(
  m = metrics_df[,c("class", "n", "P", "PPV", "bsp_id", "class1")],
  metric_title = "Positive predictive value (PPV)",
  metric_shrt = "PPV [-]"
)
```

\newpage

\section*{Accuracy}

Accuracy is a metric used in classification to measure the proportion of correctly predicted instances. It indicates how often a model’s predictions match the true labels. For a dataset with \(N\) samples, where \(TP\) = true positives, \(TN\) = true negatives, \(FP\) = false positives, and \(FN\) = false negatives, accuracy is:

\[
\text{Accuracy} = \frac{TP + TN}{TP + TN + FP + FN}
\]

It ranges from 0 to 1, with 1 being perfect prediction. While intuitive, accuracy can be misleading for imbalanced datasets.


```{r Accuracy, echo=FALSE, fig.height=7}
metric_plot(
  m = metrics_df[,c("class", "n", "P", "accuracy", "bsp_id", "class1")],
  metric_title = "Accuracy",
  metric_shrt = "Accuracy [-]"
)
```

\newpage

```{r BSPIDs, echo=FALSE, out.width="\\textwidth", fig.width=13, fig.height=5.9}
# Plot model objects generated in the run directory.
rf_files = params$rf_files # paste0("../", )

par(mfrow = c(length(rf_files), 1), mar = c(3, 4, 1, 1))
for (f in rf_files) {
  plot(damageRF::plotter(readRDS(f)))
}
```

\newpage

# {damageRF} Dependencies

```{r DepPlot, echo=FALSE, out.width="\\textwidth", fig.width=7, fig.height=7}
# Draw dependency graph for the `{damageRF}` package.
par(mar = c(0,0,0,0))

dep = suppressMessages({damageRF::dependency_report()})

plot(
  dep$tree$graph,
  layout = dep$tree$coords,
  vertex.label = dep$tree$labels,
  vertex.label.cex = 0.3,
  vertex.label.color = "black",
  vertex.size = 7,
  edge.arrow.size = 0.2,
  vertex.color = dep$tree$colors
)
```
