---
title: "run_rf2"
author: "JW"
date: "`r Sys.Date()`"
output: html_document
---

## Setup

```{r Run}
# docker run --rm \
#     -v /app/:/app \
#     -v /data/:/data \
#     -v /output/:/output \
#     -e DATASET_PATH="/data/datasets/$filename" \
#     justusweyers/04randomforest_dgx
```

docker run --rm -v C:/Users/justu/Desktop/damagedata_analyser/R/04randomforest_dgx/:/app -v C:/Users/justu/Desktop/damagedata_analyser/R/04randomforest_dgx/:/data -v C:/Users/justu/Desktop/damagedata_analyser/R/04randomforest_dgx/:/output -e DATASET_PATH="/data/datasets/$filename" justusweyers/04randomforest_dgx


#### Package

```{r Packages}
# Install local dependency only when not already available.
require(ranger)

if (!requireNamespace("damageRF", quietly = TRUE)) {
  install.packages("remotes")
  remotes::install_github("JustusWeyers/damagedata_analyser/R/packages/damageRF")
}
```

#### Helpers

```{r rf2Helpers}
# Reuse plotting and PDF export helpers.
source("helpers.R")
```

## Training

#### Training characteristics

```{r TrainingCharacteristics}
# Store run metadata and outputs for reporting.
training_characteristics = list(
  errors = list(),
  wd = getwd(),
  data_dir = "/data/datasets/",
  output_dir = "/output/"
)

training_characteristics[["session_info"]] = sessionInfo()
training_characteristics[["model_name"]] = "rf2"
```

#### Loop

Iterate through datasets. 

```{r Loop}

db_path = Sys.getenv("DATASET_PATH")

# Take timestamp
training_characteristics[["start"]] = format(Sys.time(), "%H:%M:%S %d.%m.%Y")

# Database connection
print("DB at:")
print(db_path)

if (!RSQLite::dbCanConnect(RSQLite::SQLite(), db_path)) {
  stop("Cannot connect to database at: ", db_path)
} else {
  db = RSQLite::dbConnect(RSQLite::SQLite(), db_path)
  print("Successfully connected to database.")
} 

# Dataset name
dataset_name = tools::file_path_sans_ext(basename(db_path))
training_characteristics[["dataset"]] = dataset_name

# Dataset output dir
training_characteristics[["dataset_out_dir"]] = paste0(
  training_characteristics[["output_dir"]], 
  "/",
  training_characteristics[["model_name"]], 
  "/",
  training_characteristics[["model_name"]], 
  "_", 
  format(Sys.time(), "%Y-%m-%d_%H-%M-%S"), 
  "/", 
  dataset_name, 
  "/"
)

if (!dir.exists(training_characteristics[["dataset_out_dir"]])) {
  dir.create(training_characteristics[["dataset_out_dir"]], recursive = TRUE)
}

# Read data
print("Reading data")
readin = DBI::dbReadTable(db, DBI::dbListTables(db)[1]) |>
  dplyr::mutate_if(is.character, factor)

training_characteristics[["skim"]] = readin |> skimr::skim()

# Get unique bsp_ids sorted by frequency
print("Extracting unique bsp_id")
bspids = names(sort(table(readin$bsp_id), decreasing = TRUE))
print(paste0("Found ", length(bspids), " unique bsp_ids."))

# Fit one random forest per BSP id and persist artifacts.
print("Starting RF model fitting")
for (bspid in bspids) {

  print(paste("Processing bsp_id:", bspid))

  log_file = file.path(training_characteristics[["dataset_out_dir"]], paste0("rf_", bspid, ".log"))
  log_con = file(log_file, open = "wt")

  sink(log_con)
  sink(log_con, type = "message")

  cat(bspid, "\n")
  cat("start: ", format(Sys.time(), "%Y-%m-%d_%H-%M-%S"), "\n\n")
  flush.console()

  tryCatch({
    rf = damageRF::rf2(bspid, prep = readin, target = "bewert_d", n = NA, trees = 1000, n_boots = 30)
    saveRDS(rf, file.path(training_characteristics[["dataset_out_dir"]], paste0("rf_", bspid, ".rds")))
    
    training_characteristics[["metrics"]][[bspid]] = rf@metrics
    training_characteristics[["size"]][[bspid]] = rf@size
    training_characteristics[["plots"]][[bspid]] = plotter(rf)
    
    flush.console()
  }, error = function(e) {
    training_characteristics[["errors"]][[bspid]] = e 
    message("Fehler bei BSP-ID ", bspid, ": ", e$message)
  })

  cat("end: ", format(Sys.time(), "%Y-%m-%d_%H-%M-%S"), "\n")
  flush.console()

  sink(type = "message")
  sink()

  close(log_con)
}

training_characteristics[["end"]] = format(Sys.time(), "%H:%M:%S %d.%m.%Y")

# Training Report

print("Report")

# Collect saved RF model files for the final report plots.
files = list.files(training_characteristics[["dataset_out_dir"]])
rf_files = startsWith(files, "rf_") & endsWith(files, ".rds")
rf_files = list.files(training_characteristics[["dataset_out_dir"]], full.names = TRUE)[rf_files]
training_characteristics[["rf_files"]] = rf_files

print(rf_files)

rmarkdown::render(
  input = "report.Rmd",
  output_format = "pdf_document",
  output_file = "report.pdf",
  output_dir = training_characteristics[["dataset_out_dir"]], 
  params = training_characteristics,
  envir = new.env()
)

saveRDS(training_characteristics, file.path(training_characteristics[["dataset_out_dir"]], "training_characteristics.rds"))

# Disconnect from database
RSQLite::dbDisconnect(db)
```

